### 2013-04-28 Documents/ms.tex
Writing

\documentclass[letterpaper, 11pt]{article}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[left=3cm,top=3cm,right=3cm]{geometry}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\parindent=0cm

\title{Document Title}
\author{Brendon J. Brewer}

\begin{document}
\maketitle

\section{Introduction}

\section{Spatial Distribution}
Suppose the two uniform variables are $u_1$ and $u_2$.
These get turned into an elliptical thing by the following
one-to-one transformation:
\begin{eqnarray}
x &=& x_c + \frac{\mathcal{R}(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)\\
y &=& y_c + \mathcal{R}(u_1)\sqrt{q}\sin\left(2\pi u_2\right)
\end{eqnarray}
where $\mathcal{R}(u)$ is some function that determines the
radial profile. The probability density for $x$ and $y$ is given by:
\begin{eqnarray}
p(x, y)
&=& \frac{p(u_1, u_2)}{\left|\det \mathbf{J}\right|}
\end{eqnarray}
where $\mathbf{J}$ is the Jacobian of the transformation:
\begin{eqnarray}
\mathbf{J} &=&
\left(
\begin{array}{cc}
\frac{\partial x}{\partial u_1} & \frac{\partial x}{\partial u_2}\\
\frac{\partial y}{\partial u_1} & \frac{\partial y}{\partial u_2}
\end{array}
\right)
\end{eqnarray}
Let's calculate the terms. Here's the first one:
\begin{eqnarray}
\frac{\partial x}{\partial u_1} &=& \frac{\mathcal{R}'(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)
\end{eqnarray}
and the second:
\begin{eqnarray}
\frac{\partial x}{\partial u_2} &=& \frac{\mathcal{R}(u_1)}{\sqrt{q}}\cos\left(2\pi u_2\right)
\end{eqnarray}

\end{document}