### 2014-08-04 Documents/Talk/ms.tex
Started writing slides

\documentclass{beamer}
\usetheme[pageofpages=of,% String used between the current page and the
                         % total page count.
          bullet=circle,% Use circles instead of squares for bullets.
          titleline=true,% Show a line below the frame title.
          alternativetitlepage=true,% Use the fancy title page.
       %   titlepagelogo=logo-polito,% Logo for the first page.
       %   watermark=watermark-polito,% Watermark used in every page.
       %   watermarkheight=100px,% Height of the watermark.
       %   watermarkheightmult=4,% The watermark image is 4 times bigger
                                % than watermarkheight.
          ]{Torino}

\author{Brendon J. Brewer}
\title{Trans-dimensional Bayesian Inference and Nested Sampling}
\institute{Department of Statistics, The University of Auckland}
\date{August 2014}

\begin{document}
\include{content}
\end{document}